---
title: "NYC Crime and Real Estate Data Project"
subtitle: "Final Tutorial - ENGM 182"
author: Vikhyat Khare, Omkar Kshirsagar, Carter Noordsij, John Sullivan
date: June 9, 2020
output:
  html_document:
    theme: paper
    number_sections: true
    highlight: kate
    toc: true
---


## Importing the Real Estate Data
The yearly real estate data we downloaded from the [city of New York's OpenData portal](https://opendata.cityofnewyork.us) were formatted in Excel files (with the exception of that from 2016-2017, which were contained in a single CSV file, and 2018, which were missing entirely). Each Excel file contained all the real estate sales in a particular borough during a particular calendar year from 2003 to 2020.

To import the Excel file data into R, we start by creating a list of the all the filenames:
```{r eval = FALSE}
# Get list of all excel file names for 2003-2015 and 2019-2020 data
all_files = c(
    list.files(path="../Data/RE_2003-2015", pattern="*.xls", full.names=TRUE, recursive=FALSE),
    list.files(path="../Data/RE_Apr2019-Mar2020", pattern="*.xls", full.names=TRUE, recursive=FALSE)
)
```
<br/>
Next, we create a custom function to pass to `lapply()` that will import the data from each file into a dataframe, and store all those dataframes into a corresponding list called `dfs_list`. Although all the data across every file were consistently formatted with the same column headers, frustratingly, they did not always start on the same row (but always started on one of the first 10 rows). Because of this, we had to do two reads of each file—an initial read of the first 10 rows and columns just to find the index of the first column header (titled "BOROUGH" in our case), and a second read to actually import the data starting from that index. [This stack overflow post](https://stackoverflow.com/questions/43242467/reading-excel-in-r-how-to-find-the-start-cell-in-messy-spreadsheets?rq=1) was very helpful in setting up this 'double-read' functionality. We also use `clean_names()` from the `janitor` package to make the column names easier to work with later on.
```{r eval = FALSE}
library(tidyverse)
library(readxl)
library(data.table)
library(janitor)

# Import data from all files into a list of datarames
dfs_list = lapply(all_files, function(filename) {
  
    # Setting up the first read
    temp_read = suppressMessages(read_excel(filename))
    desired_sheet = 1
    skip_rows = NULL
    col_skip = 0
    search_string = "BOROUGH"
    max_cols_to_search = 10
    max_rows_to_search = 10
    
    # First read to get the cell index of the first column header
    while (length(skip_rows) == 0) {
        col_skip = col_skip + 1
        if (col_skip == max_cols_to_search) break
        skip_rows = which(stringr::str_detect(temp_read[1:max_rows_to_search,col_skip][[1]],search_string)) - 0
    }
  
    # Second read to actually import the data
    df = suppressMessages(read_excel(
        filename,
        sheet = desired_sheet,
        skip = skip_rows,
        .name_repair="universal"
    ) ) %>% clean_names()
  
    return(df)
})
```
<br/>
The only data we still need to import are the 2016-2017 data from the CSV file. We append it to our list of dataframes before combining that list into a single dataframe with `rbindlist()` and saving it as in an RDA file for easy access later on.
```{r eval = FALSE}
# Append 2016-2017 data from csv file to list of data frames
dfs_list = append(dfs_list, list(read_csv("../Data/RE_2016-2017/nyc-rolling-sales.csv")[,-(1)] %>% clean_names()))

# Combine list of data frames into a single data frame and save as RDA file
df_housing = rbindlist(dfs_list)
save(df_housing, file="Data_Housing.rda")
```
<br/>
Now that we have our housing data, we can move on to importing the crime data!
<br/>
```{r echo = FALSE, message = FALSE, warning = FALSE}
setwd("/Users/carternoordsij/Documents/dartmouth/Academics/Class materials/engm_182/Engm182_Project")
load("Data_Housing.rda")
library(knitr)
library(kableExtra)

kable(head(df_housing)) %>%
    kable_styling() %>%
    scroll_box(width = "100%", height = "300px")
```
<br/>
