---
title: "NYC Crime and Real Estate Data Project"
subtitle: "Final Tutorial - ENGM 182"
author: Vikhyat Khare, Omkar Kshirsagar, Carter Noordsij, John Sullivan
date: June 9, 2020
output:
  html_document:
    number_sections: true
---

# Downloading the census block data
Download the United States census bureau data estimates for 2015 for all 2,167 individual NYC census tracts.
This data is in two parts. Import the first part which includes the block code/census tract, its latitude and longitude as well as the name of the county and state through the following code. After getting them in a dataframe (block_loc), remove the rows with 'NA'. A basic 'unique value check' shows that there are 38,936 locations in 20,406 blocks in 15 counties. These longitude and latitude locations would help us later when we classify the demographic data by zip code.

```{r eval = FALSE}
#Code for census data
block_loc <- read.csv("../Data/Census_2015/census_block_loc.csv")
head(block_loc)
summary(block_loc)
nrow(block_loc)
#This data frame has 5 columns (Latitude, longitude and block code for a block as well as its county and state) and 38,396 rows


#Checking if there are any missing or infinite 
missing_val = any(is.na(block_loc))
#There are no missing values in the dataframe

#Checking for unique values 
unique_state <- unique(block_loc$State)
unique_county <- unique(block_loc$County)
unique_block <- unique(block_loc$BlockCode)
#There are 38,936 locations in 20,406 blocks which are in 15 counties in NJ and NY in the data
```
<br/>
The dataframe looks like this:
```{r echo = FALSE, include= FALSE}
library(knitr)
library(kableExtra)
kable(head(block_loc))%>%
    kable_styling()%>%
    scroll_box(width = "100%", height = "300px")
```
<br/>
We divide this dataframe into different counties in NY and NJ to get a better sense about the data. This gives us information about the number of locations in each of the counties in our data. 

```{r eval = FALSE}
#Splitting dataframes into different states and then different counties
library(dplyr)
block_loc_NY <- filter(block_loc, State == "NY")
#This has 24,240 locations
block_loc_NJ <- filter(block_loc, State == "NJ")
#This has 14,156 locations
unique(block_loc_NY$County)
block_loc_Richmond <- filter(block_loc, County == "Richmond") #3933 locations
block_loc_Queens <- filter(block_loc, County == "Queens") #6872 locations
block_loc_Nassau <- filter(block_loc, County == "Nassau") #4906 locations
block_loc_Kings <- filter(block_loc, County == "Kings") #3729 locations
block_loc_NewYork <- filter(block_loc, County == "New York") #1295 locations
block_loc_Bronx <- filter(block_loc, County == "Bronx") #2224 locations
block_loc_Westchester <- filter(block_loc, County == "Westchester") #1281 locations

unique(block_loc_NJ$County)
block_loc_Middlesex <- filter(block_loc, County == "Middlesex") #760 locations
block_loc_Monmouth <- filter(block_loc, County == "Monmouth") #1117 locations
block_loc_Union<- filter(block_loc, County == "Union") #1553 locations
block_loc_Hudson <- filter(block_loc, County == "Hudson") #2400 locations
block_loc_Essex <- filter(block_loc, County == "Essex") #3302 locations
block_loc_Bergen <- filter(block_loc, County == "Bergen") #3456 locations
block_loc_Passaic <- filter(block_loc, County == "Passaic") #1539 locations
block_loc_Morris <- filter(block_loc, County == "Morris") #29 locations
```
<br/>
# Downloading the 2nd part NYC census tracts data
The second part of the census data is the NYC census tracts data which is imported below. This is also split into different dataframes which contained data about sex, race, income, jobs, commute and employment, albeit these dataframes will not be used separately for analysis in the future. 
```{r eval = FALSE}
#importing nyc census tracts file
nyc_tracts <- read.csv("../Data/Census_2015/nyc_census_tracts.csv")
head(nyc_tracts)

#Splitting this into different dataframes
nyc_tracts_sex <- select(nyc_tracts, CensusTract, County, Borough, TotalPop, Men, Women) #Actual number of men and women in population
nyc_tracts_race <- select(nyc_tracts, CensusTract, County, Borough, TotalPop, Hispanic, White, Black, Native, Asian) #Percentage of different race in population
nyc_tracts_income <- select(nyc_tracts, CensusTract, County, Borough, TotalPop, Citizen, Professional, Service, Office, Construction, Production) #Percentage of different jobs
nyc_tracts_jobs <- select(nyc_tracts, CensusTract, County, Borough, TotalPop, Citizen, Income, IncomeErr, IncomePerCap, IncomePerCapErr, Poverty, ChildPoverty) #Income and Poverty data
nyc_tracts_commute <- select(nyc_tracts, CensusTract, County, Borough, TotalPop, Citizen, Drive, Carpool, Transit, Walk, OtherTransp, WorkAtHome, MeanCommute) #Percentage of different ways that people use for commuting. Don't know MeanCommute, maybe it is the percentage of people commuting every day
nyc_tracts_employment <- select(nyc_tracts, CensusTract, County, Borough, TotalPop, Citizen, Employed, PrivateWork, PublicWork, SelfEmployed, FamilyWork, Unemployment)#Percentage of different kinds of employment
```
<br/>
The second dataframe looks like this:
```{r echo = FALSE, include= FALSE}
library(knitr)
library(kableExtra)
kable(head(nyc_tracts))%>%
    kable_styling()%>%
    scroll_box(width = "100%", height = "300px")
```