---
title: "NYC Crime and Real Estate Data Project"
subtitle: "Final Tutorial - ENGM 182"
author: Vikhyat Khare, Omkar Kshirsagar, Carter Noordsij, John Sullivan
date: June 9, 2020
output:
  html_document:
    number_sections: true
---
# Adding zipcode to each criminal activity
The df_crime dataframe that contains data about all the crimes committed is first of all purged of all NA rows in the latitude and longitude columns. Then by reverse geocoding, the latitude and longitude values are mapped to a zip code in the following way.

```{r eval = FALSE}
# These should be taken care of with the higher level Import_All.R:
# source("../Engm182_Project/Import_Scripts/Import_Crime_Data.R")
# head(df_crime)
library(tidyr)
library(dplyr)
#Dropping the rows whcih have blank in latitude or longitude column
df_crime2 <- drop_na(df_crime, Latitude, Longitude)

#Reverse geocoding to find the zip code for each Lonlat
library("ggmap")
dset_crime2 <- as.data.frame(df_crime2[,29:28]) 
loc <- dset_crime2
res <- lapply(seq(nrow(location)), function(i){ 
    revgeocode(location[i,], 
               output = c("all"), 
               messaging = FALSE, 
               sensor = FALSE, 
               override_limit = FALSE)
})
```

# Assigning weights to each crime
Assign a weight to every crime based on the degree of severity. A violation which includes minor offenses like speeding ticket, public intoxication etc. is assigned a weight of 1. A misdemeanour which can include a simple assault or driving under influence is assigned a wight of 2. These were easy to identify as the LAW_CAT_CD column in df_crime gives the category of the crime (Violation, Misdemeanour or Felony).

```{r eval = FALSE}
#Assigning weights to each crime
#Violations are the most minor of offenses. They have been assigned a weight of 1
#source: https://patrickparrottalaw.com/differences-between-a-violation-a-misdemeanor-and-a-felony/
df_crime2$weight <- 0
df_crime2$weight[which(df_crime2$LAW_CAT_CD=="VIOLATION")]=1

#Misdemeanours are more serious than violations but less severe than felonies, they can carry up to a year in jail and have score of 2
#source: https://patrickparrottalaw.com/differences-between-a-violation-a-misdemeanor-and-a-felony/
df_crime2$weight[which(df_crime2$LAW_CAT_CD=="MISDEMEANOR")]=2
```
<br/>
Assigning weights to felonies is a little more tricky. Since, felonies include a vast variety of criminal activity, we decided to first divide them into 4 categories, 1st to 4th degree. First-degree felonyincludes murder, rape, kidnapping, arson and fraud. Second-degree felony includes aggravated assault, felony assault, arson, manslaughter, possession of a controlled substance and child molestation. Third-degree felony includes assault and battery, elder abuse, transmission of pornography, driving under the influence, fraud and arson. Finally, fourth-degree felony includes involuntary manslaughter, burglary, larceny and resisting arrest. (These lists are just examples and not exhaustive and are based on a source: https://legaldictionary.net/felony/) From the OFNS_DESC column generate a list of unique keywords and group them into 4 lists pertaining to the type of feloniy. Then, the crime is assigned a weight of 6, 5, 4 or 3 for felonies of 1st degree, 2nd degree, 3rd degree and 4th degree respectively depending on which category the OFNS_DESC column of that crime suggests.
```{r eval = FALSE}
#Felonies are the most serious of offenses and will be sub-classified
#Common law and statutes in most states divide felonies into first through fourth degree felonies, each carrying decreasing penalties
#source: https://legaldictionary.net/felony/

df_crime2_felony <- subset(df_crime2, df_crime2$LAW_CAT_CD=="FELONY")
felonytypes <- unique(df_crime2_felony$OFNS_DESC)
felony_1stdegree <- c("RAPE", "THEFT-FRAUD", "ARSON", "KIDNAPPING & RELATED OFFENSES", "MURDER & NON-NEGL. MANSLAUGHTER", "KIDNAPPING", "KIDNAPPING AND RELATED OFFENSES", "NYS LAWS-UNCLASSIFIED FELONY" )
felony_2nddegree <- c("FELONY ASSAULT", "FORGERY", "SEX CRIMES", "DANGEROUS DRUGS", "DANGEROUS WEAPONS")
felony_3rddegree <- c("INTOXICATED/IMPAIRED DRIVING","ROBBERY", "CRIMINAL MISCHIEF & RELATED OF", "PROSTITUTION & RELATED OFFENSES")
felony_4thdegree <- c("GAMBLING","", "BURGLARY","GRAND LARCENY", "GRAND LARCENY OF MOTOR VEHICLE", "MISCELLANEOUS PENAL LAW", "POSSESSION OF STOLEN PROPERTY", "CHILD ABANDONMENT/NON SUPPORT", "HOMICIDE-NEGLIGENT-VEHICLE", "HOMICIDE-NEGLIGENT,UNCLASSIFIE", "ENDAN WELFARE INCOMP")

df_crime2$weight[which(df_crime2$OFNS_DESC %in% felony_1stdegree == TRUE)]=6
df_crime2$weight[which(df_crime2$OFNS_DESC %in% felony_2nddegree == TRUE)]=5
df_crime2$weight[which(df_crime2$OFNS_DESC %in% felony_3rddegree == TRUE)]=4
df_crime2$weight[which(df_crime2$OFNS_DESC %in% felony_4thdegree == TRUE)]=3

df_crime <- df_crime2
save(df_crime, file='Data_Crime.rda')

#Extract month and year for each crime
df_crime3 <- subset(df_crime2, select = c(CMPLNT_FR_DT, OFNS_DESC, LAW_CAT_CD, Longitude, Latitude, weight))
head(df_crime3)
```

