---
title: "NYC Crime and Real Estate Data Project"
subtitle: "Final Tutorial - ENGM 182"
author: Vikhyat Khare, Omkar Kshirsagar, Carter Noordsij, John Sullivan
date: June 9, 2020
output:
  html_document:
    number_sections: true
---

# Group crime information by month, year, and zipcode

First, the relevant datafiles are imported to the environment and the packages are loaded:

```{r eval=FALSE}
library(lubridate)
library(dplyr)
load(file = "Data_Crime_w_Zipcodes.rda")
load(file = "Data_Zip_Population.rda")
```

For analysis, it is helpful to have the events grouped by month and year, then by a given zipcode. Then, a field like the crime score can be aggregated an normalized for this criteria. In this case, the crime data is formatted in the desired way with the following command:

```{r eval=FALSE}
score_by_time_and_rating <- df_crime %>% group_by(month=floor_date(CMPLNT_FR_DT, "month"), zip_code)  %>% summarize(summary_variable=sum(weight))
```

Source: https://stackoverflow.com/questions/33221425/how-do-i-group-my-date-variable-into-month-year-in-r

## Group other fields by time and zipcode

In addition, commands were run to normalize the crime score by the total population and only take data greater than the year 2000. There was extraneous data for crimes hundreds of years in the past and 2000 was chosen as a good baseline. It also appeared that the reporting method for these crimes changes in 2006, as evidenced by a dramatic change in the crime scores at this time.

```{r class.source = NULL, eval = FALSE}
# Summing the scores
names(score_by_time_and_rating)[names(score_by_time_and_rating) == 'summary_variable'] <- 'sum_weight'

# Joining the zipcodes
score_by_time_and_rating <- left_join(score_by_time_and_rating, zip_population, by="zip_code")

# Dividing sum by the total population (at 2015 census levels)
score_by_time_and_rating$weight <- score_by_time_and_rating$sum_weight / score_by_time_and_rating$TotalPop

# Dipslaying the head
head(score_by_time_and_rating)

save(score_by_time_and_rating, file = "Data_Score_by_Time_and_Rating.rda")

#Group by year and rating (This is for the regression model for house sales)
score_by_zip_and_year <- df_crime %>% group_by(year=floor_date(CMPLNT_FR_DT, "year"), zip_code)  %>% summarize(summary_variable=sum(weight))

#Summing the scores
names(score_by_zip_and_year)[names(score_by_zip_and_year) == 'summary_variable'] <- 'sum_weight'

#Extracting the year
score_by_zip_and_year$year <- year(score_by_zip_and_year$year)
score_by_zip_and_year <- subset(score_by_zip_and_year, select = c(year,zip_code,sum_weight))

# Joining the zipcodes with populations
score_by_zip_and_year <- left_join(score_by_zip_and_year, zip_population, by="zip_code")
score_by_zip_and_year <- subset(score_by_zip_and_year, select = c(year,zip_code,sum_weight, TotalPop))

# Dividing sum by the total population (at 2015 census levels)
score_by_zip_and_year$weight <- score_by_zip_and_year$sum_weight / score_by_zip_and_year$TotalPop

#Taking data only for years 2001 and above 
score_by_zip_and_year <- subset(score_by_zip_and_year, year > 2000)
score_by_zip_and_year <- subset(score_by_zip_and_year, select = c(zip_code, year, sum_weight, weight, TotalPop))

save(score_by_zip_and_year, file = "Data_Score_by_year_and_zipcode.rda")
```
