---
title: "NYC Crime and Real Estate Data Project"
subtitle: "Final Tutorial - ENGM 182"
author: Vikhyat Khare, Omkar Kshirsagar, Carter Noordsij, John Sullivan
date: June 9, 2020
output:
  html_document:
    number_sections: true
---

# Converting coordinates to zipcodes

There was difficulty with initially processing the location data into a format for analysis. For the crime and census data:

* Crime data - complaints were initially recorded with only a latitude and longitude
* Census data - the census data used census tracts for identifying locations

For the regressions and displaying the data visually, the data needs to be grouped by similar location. In this case, zipcode was chosen to be the level to group by. By classifying the data entries in a zipcode, the effect of the zipcode can be studied.

At first, an API was used to convert the coordinates from the crime data to zip codes. The revgeocode function from the ggmap package was used and applied to each crime complaint. However, it was taking too long to make the requests to retrieve each of the zipcodes. Although this route was not pursued, it may be possible to perform the API requests in parallel.

Instead, it was chosen to perform the reverse geocoding locally, as to avoid having to make requests to an API service for each zip code. Further, since the zip codes were limited to only New York, it was possible to only download the location data for this area.

## Sources for latitude and longitude to zipcode code and data

The source for the code to retrieve the zipcodes from latitude and longitude coordinates locally is: https://stackoverflow.com/questions/46267287/reverse-geocoding-speed

The source for the geospatial polygon data to process the coordinates is: https://jsspina.carto.com/tables/nyc_zip_code_tabulation_areas_polygons/public/map

## Processing the crime data

First, the packages are loaded:

```{r eval=FALSE}
library(sf)
library(purrr)
library(lubridate)
library(dplyr)
```


The crime data is read in with the following command:

```{r eval=FALSE}
sf <- sf::st_read("nyc_zip_code_tabulation_areas_polygons.geojson")
```

Then the desired latitude and longitude fields are taken. the complaint number is also taken so that the crime-zipcode data can be joined again later. The NA values are omitted.

```{r eval=FALSE}
CMPLNT_NUM <- c(df_crime$CMPLNT_NUM)
latitude <- c(df_crime$Latitude)
longitude <- c(df_crime$Longitude)

x = data.frame(CMPLNT_NUM, longitude, latitude)
x <- na.omit(x)

sf_x <- sf::st_as_sf(x, coords = c("longitude", "latitude"))
```

As stated in the sourced guide above for reverse geocoding this data, "To perform spatial operations, the coordinate reference system needs to match between the two geometries." 

```{r eval=FALSE}
st_crs(sf_x) <- st_crs(sf)
```

The coordinates are processed with the st_within function and put into an output using sapply:

```{r eval=FALSE}
res <- st_within(sf_x, sf)

out <- sapply(res, function(x) as.character(sf$postalcode[x]))
```

The output for each zipcode was a list of zipcodes even though there was just one. There were multiple outputs in the case that one was trying to link overlapping geometries with the data. The zipcodes were converted to the proper format:

```{r eval=FALSE}
out2 <- map(out, 1) # Take first element from each list
out3 <- lapply(out, function(x) ifelse(length(x) == 1, x[1], NA)) # Handle NAs
out4 <- vapply(out3, paste, collapse = ", ", character(1L)) # Flattten listst; source: https://stackoverflow.com/questions/24829027/unimplemented-type-list-when-trying-to-write-table

sf_x$zip_code <- out4
```

Finally, the zipcodes are linked back with the crime data and saved:

```{r eval=FALSE}
# Putting it back with the sf_data
df_crime <- left_join(df_crime, sf_x, by = "CMPLNT_NUM")

# Format column as dates
df_crime$CMPLNT_FR_DT <- as.Date(df_crime$CMPLNT_FR_DT, format = "%m/%d/%Y")

# This column NEEDS to be removed because it causes errors
df_crime$geometry<-NULL

# Saving the data with thet important information
save(df_crime, file='Data_Crime_w_Zipcodes.rda')
```

## Processing the census data

First, the packages are loaded:

```{r eval=FALSE}
library(readxl)
library(dplyr)
library(plyr)
```

The process for finding the zipcodes that corresponded to the census data was different. The census tract information is more granular than at the zipcode level. The mapping between tracts and zipcodes is imported and joined to the census data, so each element has a corresponding zipcode.

```{r eval=FALSE}
tract_zips <- my_data <- read_excel("TRACT_ZIP_032020.xlsx")

names(tract_zips)[names(tract_zips) == 'ZIP'] <- 'zip_code'
names(tract_zips)[names(tract_zips) == 'TRACT'] <- 'CensusTract'
tract_zips$CensusTract <- as.numeric(tract_zips$CensusTract)

nyc_tracts <- left_join(nyc_tracts, tract_zips, by="CensusTract")

# Save updated nyc_tracs
save(nyc_tracts, file='Data_NYC_Tracts.rda')

#Adding all the parameters required for regression in zip_population file
zip_population <- subset(nyc_tracts, select = c(CensusTract,zip_code,TotalPop, Men, Women, Hispanic, White, Black, Native, Asian, IncomePerCap, Unemployment))
```

The most important element was the corresponding populations for each zipcode. The population data for each demographic was summed across each zipcode. The demographic information was initially stored as percentages, so it was necessary to multiply those percentages by the total population for each census tract.

```{r eval=FALSE}
#Converting the demographic and unemployment percentages to numbers and income per capita to income
zip_population$Hispanic.number <- zip_population$Hispanic*zip_population$TotalPop/100
zip_population$White.number <- zip_population$White*zip_population$TotalPop/100
zip_population$Black.number <- zip_population$Black*zip_population$TotalPop/100
zip_population$Native.number <- zip_population$Native*zip_population$TotalPop/100
zip_population$Asian.number <- zip_population$Asian*zip_population$TotalPop/100
zip_population$IncomeTot <- zip_population$IncomePerCap*zip_population$TotalPop
#Percentage of people in the age group of 18-65 in NY is 65% source: https://www.census.gov/quickfacts/newyorkcitynewyork
zip_population$Unemployment.number <- zip_population$Unemployment*0.65*zip_population$TotalPop/100

#Aggregating the various census tract data into 1 zip code
zip_population <- subset(zip_population, select = c(zip_code,TotalPop, Men, Women, Hispanic.number, White.number, Black.number, Native.number, Asian.number, IncomeTot, Unemployment.number))
zip_population <- aggregate(.~zip_code, data = zip_population, FUN = sum)

#Convert back the demographics and unemployment to % and IncomeTot to per capita income
zip_population$Hispanic.number <- zip_population$Hispanic.number*100/zip_population$TotalPop
zip_population$White.number <- zip_population$White.number*100/zip_population$TotalPop
zip_population$Black.number <- zip_population$Black.number*100/zip_population$TotalPop
zip_population$Native.number <- zip_population$Native.number*100/zip_population$TotalPop
zip_population$Asian.number <- zip_population$Asian.number*100/zip_population$TotalPop
zip_population$IncomeTot <- zip_population$IncomeTot/zip_population$TotalPop
zip_population$Unemployment.number <- zip_population$Unemployment.number*100/(0.65*zip_population$TotalPop)
```

Finally, the columns were renamed as needed. The data for zipcode to population demographic was saved for analysis later:

```{r eval=FALSE}
#Renaming the column headers
colnames(zip_population) <- c("zip_code", "TotalPop", "Men", "Women", "Hispanic", "White", "Black", "Native", "Asian", "PerCapitaIncome", "Unemployed")

save(zip_population, file = "Data_Zip_Population.rda")
```

## Source for census tract to zipccode mapping data

The following link includes the updated census tract to zipcode information: https://www.huduser.gov/portal/datasets/usps_crosswalk.html