---
title: "NYC Crime and Real Estate Data Project"
subtitle: "Final Tutorial - ENGM 182"
author: Vikhyat Khare, Omkar Kshirsagar, Carter Noordsij, John Sullivan
date: June 9, 2020
output:
  html_document:
    number_sections: true
---
# Section of housing sales data
First, load the 2 relevant files. Data_Housing.rda (df_housing) contains all the housing sales data. Data_Zip_Population.rda (zip_population) contains the census data for each zip code.
```{r eval = FALSE}
load(file="Data_Housing.rda")
load(file="Data_Zip_Population.rda")
```
<br/>
Next, remove rows which are either blank or have NA values in the sale_price or land_square_feet column of df_housing. 
```{r eval = FALSE}
# Only sales prices > 0 for housing data

# Only take > 0; source: https://stackoverflow.com/questions/33535006/remove-rows-if-it-is-negative-number
df <- df_housing[df_housing$land_square_feet > 0, ]
df <- df[df$sale_price > 0, ]

# Get rid of NA values; source: https://stackoverflow.com/questions/48658832/how-to-remove-row-if-it-has-a-na-value-in-one-certain-column/48659101
df <- df[!is.na(df$sale_price), ]
df <- df[!is.na(df$land_square_feet), ]
```
<br/>
Create a new dataframe, df_sale, which containse the relevant sales price prediction input variables from the df_housing dataframe. Add 2 more column, sale_year and sale_month, which contain the year and month of the sale. These values are extracted from the sale_date column by using the functions year() and month() respectively.
<br/>
Add a new column, crime_score_year. If the sale was completed in the months of January, February or March, add the year just before the sale_year in the crime_score_year column. If the sale was completed in the rest of the months, crime_score_year = sale_year. This is done to map the relevant year's crime score for each sale's analysis. For e.g. if a sale happens in January of 2006, then it doesn't make sense to add the crime score for 2006. Crimes happening in 2005 would have affected that sale's price more than 2006.
```{r eval = FALSE}
#Identify the input variables for sales price prediction
#Zipcode (will be used to join census and crime data), land_square_feet, gross_square_feet, year built, tax_class_at_time_of_sale, building_class_at_time_of_sale, sale_date (will be used to link crime rating for that year)
#Subset new dataframe with relevant variables
df_sale <- subset(df, select = c(zip_code,land_square_feet, gross_square_feet, year_built, tax_class_at_time_of_sale, building_class_at_time_of_sale, sale_date, sale_price))
df_sale$sale_year <- year(df_sale$sale_date) 
df_sale$sale_month <- month(df_sale$sale_date)
#If sale is in months 1, 2, 3 then crime score will be taken from previous year else crime score of that year
df_sale$crime_score_year <- ifelse(df_sale$sale_month == 1 | df_sale$sale_month == 2 | df_sale$sale_month == 3, df_sale$sale_year-1, df_sale$sale_year)
```

# Adding the section about census data
Now, we will add the census data to the housing sales data. First find the zipcodes which are present in both df_sale and zip_population dataframes. Then add the df_sale and zip_population dataframes by these unique zipcodes through the following code. This merged dataframe is saved as df_sale_census.
```{r eval = FALSE}
#Adding the census data by linking through zipcode
#Taking a subset of df_sale for only the zip codes for which we have census data
zip_codes_inhousing <- unique(df_sale$zip_code)
zip_codes_incensus <- unique(zip_population$zip_code)
zip_codes_inhousing_incensus <- intersect(zip_codes_inhousing, zip_codes_incensus)

#Merging the sales and census data into one dataframe
df_sale$zip_code <- as.character(df_sale$zip_code)
df_sale_census <- merge(df_sale[df_sale$zip_code %in% zip_codes_inhousing_incensus, ], zip_population[zip_population$zip_code %in% zip_codes_inhousing_incensus, ], by = "zip_code")
```

# Adding the crime score
Load the file Data_Score_by_year_and_zipcode.rda (score_by_zip_and_year) that was created in section "04B: Data_Processing_Month_Year". Find the zipcodes that are present in both df_sale_census and score_by_zip_and_year and store them in zip_codes_inhousing_incensus_incrime. Merge the df_sale_census and score_by_zip_year dataframes by zipcodes and store them in df_sale_census_crime. Remove all rows from this final dataframe which have a zero value and save this dataframe as Data_sale_census_crime.rda for future analysis.
```{r eval = FALSE}
load(file = "Data_Score_by_year_and_zipcode.rda")
#Adding the crime score data for each year
zip_codes_inhousing_incensus_incrime <- intersect(unique(df_sale_census$zip_code), unique(score_by_zip_and_year$zip_code))
colnames(score_by_zip_and_year) <- c("zip_code", "crime_score_year", "sum_weight", "weight", "TotalPop")
df_sale_census_crime <- merge(df_sale_census[df_sale_census$zip_code %in% zip_codes_inhousing_incensus_incrime, ], score_by_zip_and_year[score_by_zip_and_year$zip_code %in% zip_codes_inhousing_incensus_incrime, ], by = c("zip_code", "crime_score_year"))
df_sale_census_crime <- na.omit(df_sale_census_crime)
#any(is.na(df_sale_census_crime))

#Remove values of 0
#source: https://stackoverflow.com/questions/9977686/how-to-remove-rows-with-any-zero-value
zero_rows = apply(df_sale_census_crime, 1, function(row) all(row != 0))
df_sale_census_crime <- df_sale_census_crime[zero_rows, ]

save(df_sale_census_crime, file = "Data_sale_census_crime.rda")
```

```{r}
library(knitr)
library(kableExtra)
kable(head(df_sale_census_crime), caption = "<p style=\"padding-left:10px\"> **Final Dataframe**</p>") %>%
    kable_styling() %>%
    scroll_box(width = "100%", height = "300px")
```