---
title: "NYC Crime and Real Estate Data Project"
subtitle: "Final Tutorial - ENGM 182"
author: Vikhyat Khare, Omkar Kshirsagar, Carter Noordsij, John Sullivan
date: June 9, 2020
output:
  html_document:
    number_sections: true
---

# Regression model

The regression model helps to establish a baseline. Unfortunately, when working with a large set of data, there are issues faced when running regressions:

1. Large dataset - it takes a long time to run the regression
2. P-values - they go to zero quickly (they all became significant)
3. Mean-squared error (MSE) - the mean square errors of the predicted values are high, despite the amount of data

Each of the issues can be respectively addressed:

1. A function called fastLm removes overhead from the regression
2. The data can be partitioned, and the accuracy prediction can be analyzed (MSE)
3. Machine learning models can be used to capture data granularity (see machine learning code section)

## Code to setup and run regressions on combined crime data

The steps for setting up the regression, running the model, and making predictions are in the following sections.

### Load packages and import data

First, the packages are loaded:

```{r eval=FALSE}
library(dplyr)
library(RcppArmadillo)
library(lubridate)
```

The desired data is imported into R for the regression.

```{r eval=FALSE}
#Regress sale price with sale year, land area, gross area, tax class, building class, year built
load(file = "Data_sale_census_crime.rda")
```

### Variable selection

The desired independent and dependent (sale_price) variables are selected with the code below and certain parameters are set to limit the amount of data.

```{r eval=FALSE}
# Make sure sale prices are numeric
df_sale_census_crime$sale_price <- as.numeric(df_sale_census_crime$sale_price)

# Remove large house sale prices; source: https://stackoverflow.com/questions/25764810/delete-rows-based-on-range-of-values-in-column
df_sale_census_crime <- df_sale_census_crime[with(df_sale_census_crime, sale_price <= 1000000), ]

#Regression 1
x <- df_sale_census_crime %>% select("land_square_feet", "PerCapitaIncome", "Unemployed", "TotalPop.x", "Men", "Women")

y <- df_sale_census_crime %>% select("sale_price")
```

### Variable coercion

The variables are coerced into the proper types for the regression (e.g. factor variable).

```{r eval=FALSE}
x$land_square_feet <- as.numeric(x[[1]])
x$PerCapitaIncome <- as.numeric(x[[2]])
x$Unemployed <- as.numeric(x[[3]])
x$TotalPop.x <- as.numeric(x[[4]])
x$Men <- as.numeric(x[[5]])
x$Women <- as.numeric(x[[6]])

# Setting up factor variable
df_sale_census_crime$sale_year <- as.character(df_sale_census_crime$sale_year)
df_sale_census_crime$year_built <- as.character(df_sale_census_crime$year_built)

building_class <- model.matrix( ~ building_class_at_time_of_sale - 1, data=df_sale_census_crime )
sale_year <- model.matrix( ~ sale_year - 1, data=df_sale_census_crime )
year_built <- model.matrix( ~ year_built - 1, data=df_sale_census_crime )

# Binding factor variables
x<-cbind(x, building_class)
x<-cbind(x, sale_year)
x<-cbind(x, year_built)

# Add intercept column and renaming
x<-cbind(x, 1)
colnames(x)[dim(x)[2]] <- "Intercept" # Source: https://www.dummies.com/programming/r/how-to-name-matrix-rows-and-columns-in-r/

# Setup the y output variable
y <- as.numeric(y[[1]])

# Display the classes
sapply(x, class)
sapply(y, class)
```

### Partition the data

The data is partitioned into training and test sets for running regressions and determining the accuracy of a model on the test set respectively. For this initial regression analysis, a subset of 80% of the data is used for training and the rest is used for testing the accuracy of the predictions.

```{r eval=FALSE}
# ----- Split data into training and validation sets ----- 
# create a list of 80% of the rows in the original dataset we can use for training
validation_index <- sort(sample(nrow(x), nrow(x)*.8))
# select 20% of the data for validation
x_test <- x[-validation_index,]
y_test <- y[-validation_index]
# use the remaining 80% of data to training and testing the models
x_train <- x[validation_index,]
y_train <- y[validation_index]
```

### Run the regression using fastLm

The regression is run using fastLm to expedite the process and remove the overhead from the normal regression functions.

```{r eval=FALSE}
# Run regression
fit1 <- fastLm(x_train, y_train)
```

### Summarize results

The results are summarized for analysis.

```{r eval=FALSE}
summary(fit1)
#R-squared value is 25.63%

# Perform predictions on test set
reg_pred1 <- predict(fit1, as.matrix(x_test))

# See the comparison
results <- data.frame(cbind(y_test, reg_pred1))
results$diff <- results$reg_pred1 - results$y_test
results$diff_squared <- '^' (results$diff, 2)
# RMS = 
rms = sqrt(mean(results$diff_squared))
rms

# RMS Error = 
rms_error = rms/mean(results$y_test)
rms_error
```

## Results from the regression models

The results that were found when the regression was run for this project are summarized:

| Measure                          | Value     |
|----------------------------------|-----------|
| Adjusted R-squared value ($R^2$) | 27.26%    |
| Root Mean Squared (RMS)          | 198,239.1 |
| Root Mean Squared Error (RMSE)   | 43.59%    |

The more detailed output from the regression model is summarized below. While it may be possible to improve the regression results by running further regressions, the machine learning models explored next will likely help to capture the granularity in the data.

<details>
<summary>Regression output</summary>
```
> # Run regression
> fit1 <- fastLm(x_train, y_train)
> summary(fit1)

Call:
fastLm.default(X = x_train, y = y_train)

Residuals:
     Min.   1st Qu.    Median   3rd Qu.      Max. 
-951240.0  -99101.0    2810.9  110820.0 1088100.0 

                                    Estimate      StdErr     t.value   p.value    
land_square_feet                  2.8982e-02  7.3239e-03  3.9572e+00 7.583e-05 ***
PerCapitaIncome                   1.5932e+00  5.5788e-02  2.8558e+01 < 2.2e-16 ***
Unemployed                       -1.2420e+04  1.6778e+02 -7.4027e+01 < 2.2e-16 ***
TotalPop.x                       -6.1569e+09  2.5369e-01 -2.4269e+10 < 2.2e-16 ***
Men                               6.1569e+09  2.4254e-01  2.5385e+10 < 2.2e-16 ***
Women                             6.1569e+09  2.9053e-01  2.1192e+10 < 2.2e-16 ***
building_class_at_time_of_saleA0  1.3952e+05  3.7534e+03  3.7171e+01 < 2.2e-16 ***
building_class_at_time_of_saleA1  1.0504e+05  1.4056e+03  7.4727e+01 < 2.2e-16 ***
building_class_at_time_of_saleA2  6.4862e+04  1.7563e+03  3.6931e+01 < 2.2e-16 ***
building_class_at_time_of_saleA3  3.3492e+05  5.4030e+03  6.1988e+01 < 2.2e-16 ***
building_class_at_time_of_saleA4  1.9749e+05  6.5229e+03  3.0276e+01 < 2.2e-16 ***
building_class_at_time_of_saleA5  3.9112e+04  1.4470e+03  2.7030e+01 < 2.2e-16 ***
building_class_at_time_of_saleA6 -1.5064e+05  7.0867e+03 -2.1257e+01 < 2.2e-16 ***
building_class_at_time_of_saleA7  1.8340e+05  5.3564e+01  3.4239e+03 < 2.2e-16 ***
building_class_at_time_of_saleA8 -1.1559e+05  1.4973e+01 -7.7196e+03 < 2.2e-16 ***
building_class_at_time_of_saleA9  6.3681e+04  2.0931e+03  3.0424e+01 < 2.2e-16 ***
building_class_at_time_of_saleB1  1.7355e+05  1.5154e+03  1.1453e+02 < 2.2e-16 ***
building_class_at_time_of_saleB2  1.5230e+05  1.4673e+03  1.0380e+02 < 2.2e-16 ***
building_class_at_time_of_saleB3  1.3706e+05  1.5737e+03  8.7097e+01 < 2.2e-16 ***
building_class_at_time_of_saleB9  1.2594e+05  2.0994e+03  5.9986e+01 < 2.2e-16 ***
building_class_at_time_of_saleC0  2.1164e+05  1.5671e+03  1.3505e+02 < 2.2e-16 ***
building_class_at_time_of_saleC1  2.3698e+05  4.3097e+03  5.4987e+01 < 2.2e-16 ***
building_class_at_time_of_saleC2  1.9560e+05  2.9009e+03  6.7429e+01 < 2.2e-16 ***
building_class_at_time_of_saleC3  2.1607e+05  2.6559e+03  8.1355e+01 < 2.2e-16 ***
building_class_at_time_of_saleC4  1.3695e+05  9.0729e+03  1.5095e+01 < 2.2e-16 ***
building_class_at_time_of_saleC5  1.9362e+05  6.7181e+03  2.8821e+01 < 2.2e-16 ***
building_class_at_time_of_saleC6 -1.3938e+05  3.8236e+02 -3.6451e+02 < 2.2e-16 ***
building_class_at_time_of_saleC7  1.6761e+05  7.7880e+03  2.1522e+01 < 2.2e-16 ***
building_class_at_time_of_saleC8 -3.9534e+05  3.8622e+00 -1.0236e+05 < 2.2e-16 ***
building_class_at_time_of_saleC9 -5.3525e+04  1.4023e+02 -3.8170e+02 < 2.2e-16 ***
building_class_at_time_of_saleCM -1.0373e+01          NA          NA        NA    
building_class_at_time_of_saleD0  8.2579e+04  2.5786e+01  3.2024e+03 < 2.2e-16 ***
building_class_at_time_of_saleD1 -3.0391e+04  4.1447e+02 -7.3326e+01 < 2.2e-16 ***
building_class_at_time_of_saleD2 -4.5215e+05  7.4887e+00 -6.0378e+04 < 2.2e-16 ***
building_class_at_time_of_saleD3 -1.8254e+05  9.2733e+01 -1.9684e+03 < 2.2e-16 ***
building_class_at_time_of_saleD4 -1.6205e+05  7.7115e+03 -2.1014e+01 < 2.2e-16 ***
building_class_at_time_of_saleD5  9.2455e+04  7.8430e+00  1.1788e+04 < 2.2e-16 ***
building_class_at_time_of_saleD6 -1.5591e+05  6.7311e+01 -2.3162e+03 < 2.2e-16 ***
building_class_at_time_of_saleD7 -1.0718e+05  9.0214e+01 -1.1881e+03 < 2.2e-16 ***
building_class_at_time_of_saleD8 -2.3003e+04  1.2368e+01 -1.8598e+03 < 2.2e-16 ***
building_class_at_time_of_saleD9 -1.9105e+05  7.6408e+01 -2.5004e+03 < 2.2e-16 ***
building_class_at_time_of_saleE1  1.5457e+05  4.4341e+02  3.4859e+02 < 2.2e-16 ***
building_class_at_time_of_saleE2  4.3728e+05  1.0070e+01  4.3426e+04 < 2.2e-16 ***
building_class_at_time_of_saleE3  1.4641e+05  3.0501e+02  4.8002e+02 < 2.2e-16 ***
building_class_at_time_of_saleE4  1.0907e+05  5.2813e+01  2.0653e+03 < 2.2e-16 ***
building_class_at_time_of_saleE7 -2.9949e+05  1.9951e+01 -1.5011e+04 < 2.2e-16 ***
building_class_at_time_of_saleE9  1.1786e+05  8.8237e+03  1.3357e+01 < 2.2e-16 ***
building_class_at_time_of_saleF1  1.5106e+05  1.6581e+02  9.1102e+02 < 2.2e-16 ***
building_class_at_time_of_saleF2  2.1667e+05  3.0781e+01  7.0391e+03 < 2.2e-16 ***
building_class_at_time_of_saleF4  1.6584e+05  6.3631e+02  2.6062e+02 < 2.2e-16 ***
building_class_at_time_of_saleF5  9.9344e+04  1.5017e+02  6.6155e+02 < 2.2e-16 ***
building_class_at_time_of_saleF8 -2.8220e+05  1.3758e+00 -2.0512e+05 < 2.2e-16 ***
building_class_at_time_of_saleF9  1.5692e+05  8.5776e+03  1.8295e+01 < 2.2e-16 ***
building_class_at_time_of_saleG0 -7.4920e+04  9.3074e+02 -8.0494e+01 < 2.2e-16 ***
building_class_at_time_of_saleG1  1.0360e+05  1.7836e+02  5.8084e+02 < 2.2e-16 ***
building_class_at_time_of_saleG2  1.3491e+05  7.8585e+03  1.7168e+01 < 2.2e-16 ***
building_class_at_time_of_saleG3  2.1477e+05  1.5108e+01  1.4215e+04 < 2.2e-16 ***
building_class_at_time_of_saleG4  2.3700e+05  1.2326e+02  1.9228e+03 < 2.2e-16 ***
building_class_at_time_of_saleG5  1.4077e+05  5.4218e+01  2.5963e+03 < 2.2e-16 ***
building_class_at_time_of_saleG6  1.1741e+05  4.5981e+01  2.5535e+03 < 2.2e-16 ***
building_class_at_time_of_saleG7 -1.6497e+04  4.4057e+03 -3.7444e+00 0.0001808 ***
building_class_at_time_of_saleG8  3.1372e+03  2.6921e+01  1.1653e+02 < 2.2e-16 ***
building_class_at_time_of_saleG9  1.2263e+05  7.6113e+03  1.6111e+01 < 2.2e-16 ***
building_class_at_time_of_saleGU  2.8203e+05  2.6368e+00  1.0696e+05 < 2.2e-16 ***
building_class_at_time_of_saleGW  2.0901e+05  9.7865e+00  2.1357e+04 < 2.2e-16 ***
building_class_at_time_of_saleH1 -1.7314e+00  2.4147e-12 -7.1701e+11 < 2.2e-16 ***
building_class_at_time_of_saleH2 -5.1859e+05  4.9986e+03 -1.0375e+02 < 2.2e-16 ***
building_class_at_time_of_saleH3 -5.6200e+05  8.2590e+03 -6.8046e+01 < 2.2e-16 ***
building_class_at_time_of_saleH4  1.6690e+05  1.9417e+01  8.5958e+03 < 2.2e-16 ***
building_class_at_time_of_saleH6  1.1853e+04  5.5558e+00  2.1334e+03 < 2.2e-16 ***
building_class_at_time_of_saleH7 -2.1248e+05  1.0631e+01 -1.9987e+04 < 2.2e-16 ***
building_class_at_time_of_saleH8  7.7614e-01          NA          NA        NA    
building_class_at_time_of_saleH9  2.4038e+04  2.7452e+01  8.7564e+02 < 2.2e-16 ***
building_class_at_time_of_saleHR  2.6907e+05  2.6293e+01  1.0234e+04 < 2.2e-16 ***
building_class_at_time_of_saleI1 -4.2960e+05  1.5147e+01 -2.8361e+04 < 2.2e-16 ***
building_class_at_time_of_saleI3  2.0919e+05  2.7345e+00  7.6501e+04 < 2.2e-16 ***
building_class_at_time_of_saleI4 -5.5951e+04  1.7755e+01 -3.1513e+03 < 2.2e-16 ***
building_class_at_time_of_saleI5  1.7474e+05  5.9982e+01  2.9132e+03 < 2.2e-16 ***
building_class_at_time_of_saleI6  1.1961e+05  2.4592e+01  4.8636e+03 < 2.2e-16 ***
building_class_at_time_of_saleI7  6.3836e+04  1.0070e+02  6.3393e+02 < 2.2e-16 ***
building_class_at_time_of_saleI9  8.6528e+04  4.7252e+01  1.8312e+03 < 2.2e-16 ***
building_class_at_time_of_saleJ1 -2.8717e+05  9.3349e+00 -3.0763e+04 < 2.2e-16 ***
building_class_at_time_of_saleJ2 -4.9867e+05  5.4591e+00 -9.1348e+04 < 2.2e-16 ***
building_class_at_time_of_saleJ4 -9.4846e+04  2.5902e+01 -3.6618e+03 < 2.2e-16 ***
building_class_at_time_of_saleJ6 -5.4367e+05  2.1236e+01 -2.5602e+04 < 2.2e-16 ***
building_class_at_time_of_saleJ7 -4.6257e+05  3.0157e+00 -1.5339e+05 < 2.2e-16 ***
building_class_at_time_of_saleJ9 -8.2693e-01          NA          NA        NA    
building_class_at_time_of_saleK1  1.3066e+05  5.0570e+03  2.5838e+01 < 2.2e-16 ***
building_class_at_time_of_saleK2  1.8075e+05  8.3157e+03  2.1736e+01 < 2.2e-16 ***
building_class_at_time_of_saleK3 -4.7864e+05  1.6826e+00 -2.8447e+05 < 2.2e-16 ***
building_class_at_time_of_saleK4  2.0110e+05  7.0629e+03  2.8473e+01 < 2.2e-16 ***
building_class_at_time_of_saleK5  1.5896e+05  1.5770e+02  1.0080e+03 < 2.2e-16 ***
building_class_at_time_of_saleK6  5.7383e+04  1.8493e+01  3.1029e+03 < 2.2e-16 ***
building_class_at_time_of_saleK7  1.8497e+05  4.8547e+01  3.8101e+03 < 2.2e-16 ***
building_class_at_time_of_saleK8 -3.5744e+05  2.2884e+01 -1.5620e+04 < 2.2e-16 ***
building_class_at_time_of_saleK9  1.6681e+05  8.5161e+03  1.9588e+01 < 2.2e-16 ***
building_class_at_time_of_saleL1 -3.2798e+05  8.1924e+01 -4.0035e+03 < 2.2e-16 ***
building_class_at_time_of_saleL2 -4.6219e+05  2.0444e+01 -2.2608e+04 < 2.2e-16 ***
building_class_at_time_of_saleL3 -1.0495e+00          NA          NA        NA    
building_class_at_time_of_saleL8 -7.9144e+04  1.0960e+02 -7.2209e+02 < 2.2e-16 ***
building_class_at_time_of_saleL9  4.3745e+04  3.5515e+01  1.2318e+03 < 2.2e-16 ***
building_class_at_time_of_saleM1  1.5553e+05  6.4322e+02  2.4180e+02 < 2.2e-16 ***
building_class_at_time_of_saleM2  4.7448e+05  1.3285e+00  3.5715e+05 < 2.2e-16 ***
building_class_at_time_of_saleM3  1.8350e+05  2.2997e+01  7.9795e+03 < 2.2e-16 ***
building_class_at_time_of_saleM4  3.5926e+05  1.0860e+01  3.3079e+04 < 2.2e-16 ***
building_class_at_time_of_saleM9  1.7297e+05  1.9606e+02  8.8223e+02 < 2.2e-16 ***
building_class_at_time_of_saleN1 -2.5565e+05  1.2544e+00 -2.0381e+05 < 2.2e-16 ***
building_class_at_time_of_saleN2 -1.0812e+05  3.7291e+01 -2.8993e+03 < 2.2e-16 ***
building_class_at_time_of_saleN3  3.8823e+05  1.4154e+00  2.7429e+05 < 2.2e-16 ***
building_class_at_time_of_saleN4  2.5708e+05  2.3806e+00  1.0799e+05 < 2.2e-16 ***
building_class_at_time_of_saleN9  7.9835e+04  4.0318e+01  1.9802e+03 < 2.2e-16 ***
building_class_at_time_of_saleO1  9.8493e+04  5.4205e+01  1.8170e+03 < 2.2e-16 ***
building_class_at_time_of_saleO2  5.6520e+04  8.4064e+01  6.7235e+02 < 2.2e-16 ***
building_class_at_time_of_saleO3 -2.5890e+05  7.4544e+01 -3.4732e+03 < 2.2e-16 ***
building_class_at_time_of_saleO4 -2.5858e+05  4.7554e+01 -5.4377e+03 < 2.2e-16 ***
building_class_at_time_of_saleO5  2.0678e+05  6.5123e+01  3.1753e+03 < 2.2e-16 ***
building_class_at_time_of_saleO6 -1.8394e+05  1.6759e+02 -1.0975e+03 < 2.2e-16 ***
building_class_at_time_of_saleO7  1.5411e+05  9.9111e+02  1.5549e+02 < 2.2e-16 ***
building_class_at_time_of_saleO8  1.8789e+05  2.4518e+02  7.6634e+02 < 2.2e-16 ***
building_class_at_time_of_saleO9  1.2536e+05  9.2309e+03  1.3580e+01 < 2.2e-16 ***
building_class_at_time_of_saleP1 -5.0139e+05  4.3587e+00 -1.1503e+05 < 2.2e-16 ***
building_class_at_time_of_saleP2  1.6676e+05  4.9800e+01  3.3487e+03 < 2.2e-16 ***
building_class_at_time_of_saleP3  4.0284e+05  9.5488e+00  4.2188e+04 < 2.2e-16 ***
building_class_at_time_of_saleP5  1.1493e+05  3.6320e+01  3.1643e+03 < 2.2e-16 ***
building_class_at_time_of_saleP6  4.5337e+05  1.9562e+01  2.3177e+04 < 2.2e-16 ***
building_class_at_time_of_saleP7  2.5193e+05  3.6547e+00  6.8933e+04 < 2.2e-16 ***
building_class_at_time_of_saleP8 -2.4175e+04  1.0560e+01 -2.2893e+03 < 2.2e-16 ***
building_class_at_time_of_saleP9  9.8652e+04  4.2028e+01  2.3473e+03 < 2.2e-16 ***
building_class_at_time_of_saleQ1 -6.3134e+05  2.0798e+01 -3.0355e+04 < 2.2e-16 ***
building_class_at_time_of_saleQ2 -1.9214e+04  7.9179e+00 -2.4267e+03 < 2.2e-16 ***
building_class_at_time_of_saleQ3 -2.1168e+05  3.9467e+00 -5.3634e+04 < 2.2e-16 ***
building_class_at_time_of_saleQ8  3.3311e+05  5.2588e+00  6.3344e+04 < 2.2e-16 ***
building_class_at_time_of_saleQ9  5.6611e+04  2.5971e+01  2.1798e+03 < 2.2e-16 ***
building_class_at_time_of_saleR0 -6.2189e+04  4.0096e+01 -1.5510e+03 < 2.2e-16 ***
building_class_at_time_of_saleR1 -2.1086e+05  8.5324e+00 -2.4713e+04 < 2.2e-16 ***
building_class_at_time_of_saleR4  1.0408e+05  2.3851e+00  4.3635e+04 < 2.2e-16 ***
building_class_at_time_of_saleR9 -1.3032e+05  7.4740e+01 -1.7436e+03 < 2.2e-16 ***
building_class_at_time_of_saleRR -2.9996e+05  4.0817e+01 -7.3490e+03 < 2.2e-16 ***
building_class_at_time_of_saleS0  1.3380e+05  1.0122e+02  1.3220e+03 < 2.2e-16 ***
building_class_at_time_of_saleS1  1.2850e+05  3.9465e+03  3.2560e+01 < 2.2e-16 ***
building_class_at_time_of_saleS2  1.7599e+05  2.8663e+03  6.1399e+01 < 2.2e-16 ***
building_class_at_time_of_saleS3  1.9130e+05  6.4023e+03  2.9880e+01 < 2.2e-16 ***
building_class_at_time_of_saleS4  2.2695e+05  6.7895e+03  3.3426e+01 < 2.2e-16 ***
building_class_at_time_of_saleS5  2.2889e+05  7.6276e+03  3.0009e+01 < 2.2e-16 ***
building_class_at_time_of_saleS9  2.2762e+05  6.1226e+03  3.7177e+01 < 2.2e-16 ***
building_class_at_time_of_saleT2  2.3426e+04  2.9571e+00  7.9217e+03 < 2.2e-16 ***
building_class_at_time_of_saleT9 -6.9349e+04  8.5254e+00 -8.1344e+03 < 2.2e-16 ***
building_class_at_time_of_saleU7 -1.1867e+05  3.5107e+00 -3.3803e+04 < 2.2e-16 ***
building_class_at_time_of_saleV0 -2.2715e+04  3.7162e+03 -6.1123e+00 9.828e-10 ***
building_class_at_time_of_saleV1 -6.4640e+04  7.6833e+03 -8.4130e+00 < 2.2e-16 ***
building_class_at_time_of_saleV2 -5.6667e+04  7.3045e+01 -7.7578e+02 < 2.2e-16 ***
building_class_at_time_of_saleV3 -6.8084e+04  9.0265e+01 -7.5427e+02 < 2.2e-16 ***
building_class_at_time_of_saleV5  3.5594e+05  4.0220e+00  8.8499e+04 < 2.2e-16 ***
building_class_at_time_of_saleV9 -1.1566e+05  6.3938e+01 -1.8090e+03 < 2.2e-16 ***
building_class_at_time_of_saleW1 -1.3223e+05  2.2746e+01 -5.8133e+03 < 2.2e-16 ***
building_class_at_time_of_saleW2  6.7435e+04  2.2155e+01  3.0438e+03 < 2.2e-16 ***
building_class_at_time_of_saleW3 -4.6763e+04  4.0345e+01 -1.1591e+03 < 2.2e-16 ***
building_class_at_time_of_saleW4 -7.7198e+04  2.8433e+00 -2.7151e+04 < 2.2e-16 ***
building_class_at_time_of_saleW6 -5.9114e+05  5.1179e+00 -1.1550e+05 < 2.2e-16 ***
building_class_at_time_of_saleW7 -3.1586e+05  8.8954e+00 -3.5508e+04 < 2.2e-16 ***
building_class_at_time_of_saleW8  2.2198e+05  2.7235e+01  8.1506e+03 < 2.2e-16 ***
building_class_at_time_of_saleW9  1.4626e+05  4.4200e+01  3.3089e+03 < 2.2e-16 ***
building_class_at_time_of_saleY1 -2.5078e+05  1.5060e+00 -1.6652e+05 < 2.2e-16 ***
building_class_at_time_of_saleY2 -3.4598e+05  7.8875e+00 -4.3864e+04 < 2.2e-16 ***
building_class_at_time_of_saleY4 -3.1975e+05  1.7551e+00 -1.8218e+05 < 2.2e-16 ***
building_class_at_time_of_saleY5 -4.7741e+05  3.0906e+00 -1.5447e+05 < 2.2e-16 ***
building_class_at_time_of_saleY6 -1.1746e+05  1.9593e+01 -5.9949e+03 < 2.2e-16 ***
building_class_at_time_of_saleZ0 -1.3031e+05  4.8316e+01 -2.6969e+03 < 2.2e-16 ***
building_class_at_time_of_saleZ1 -3.3428e+05  3.9923e+00 -8.3731e+04 < 2.2e-16 ***
building_class_at_time_of_saleZ2 -2.1807e+05  2.8958e+01 -7.5306e+03 < 2.2e-16 ***
building_class_at_time_of_saleZ3 -1.2179e+05  3.0810e+00 -3.9531e+04 < 2.2e-16 ***
building_class_at_time_of_saleZ8 -2.0401e+05  1.4632e+00 -1.3942e+05 < 2.2e-16 ***
building_class_at_time_of_saleZ9  4.2683e+04  7.9866e+03  5.3443e+00 9.081e-08 ***
sale_year2003                    -1.0892e+05  1.1024e+03 -9.8802e+01 < 2.2e-16 ***
sale_year2004                    -5.1748e+04  1.0476e+03 -4.9395e+01 < 2.2e-16 ***
sale_year2005                     1.2800e+04  1.0659e+03  1.2008e+01 < 2.2e-16 ***
sale_year2006                     5.6075e+04  1.1214e+03  5.0002e+01 < 2.2e-16 ***
sale_year2007                     5.4919e+04  1.2462e+03  4.4070e+01 < 2.2e-16 ***
sale_year2008                     8.8428e+03  1.3830e+03  6.3940e+00 1.618e-10 ***
sale_year2009                    -3.6373e+04  1.4479e+03 -2.5120e+01 < 2.2e-16 ***
sale_year2010                    -3.7572e+04  1.4401e+03 -2.6089e+01 < 2.2e-16 ***
sale_year2011                    -3.6054e+04  1.5219e+03 -2.3690e+01 < 2.2e-16 ***
sale_year2012                    -2.8094e+04  1.4739e+03 -1.9061e+01 < 2.2e-16 ***
sale_year2013                    -8.9054e+03  1.3906e+03 -6.4040e+00 1.515e-10 ***
sale_year2014                     1.5791e+04  1.3950e+03  1.1320e+01 < 2.2e-16 ***
sale_year2015                     3.2799e+04  1.3994e+03  2.3437e+01 < 2.2e-16 ***
sale_year2016                     6.7075e+04  2.3071e+03  2.9073e+01 < 2.2e-16 ***
sale_year2017                     7.8186e+04  1.7508e+03  4.4658e+01 < 2.2e-16 ***
sale_year2019                     1.3103e+05  1.6770e+03  7.8134e+01 < 2.2e-16 ***
sale_year2020                     1.5247e+05  3.2118e+03  4.7472e+01 < 2.2e-16 ***
year_built0                      -1.3214e+04  4.1824e+03 -3.1593e+00 0.0015815 ** 
year_built1380                   -5.1417e+05  1.2867e+00 -3.9962e+05 < 2.2e-16 ***
year_built1800                   -2.3758e+04  7.1415e+01 -3.3268e+02 < 2.2e-16 ***
year_built1814                   -1.2841e+05  2.4966e+00 -5.1435e+04 < 2.2e-16 ***
year_built1825                   -5.7477e+05  2.6333e+00 -2.1828e+05 < 2.2e-16 ***
year_built1826                    9.3972e+04  4.4656e+00  2.1044e+04 < 2.2e-16 ***
year_built1827                    7.1977e+05  4.6640e+00  1.5432e+05 < 2.2e-16 ***
year_built1838                   -4.0203e+05  3.1975e+00 -1.2573e+05 < 2.2e-16 ***
year_built1839                   -5.6975e+05  2.6778e+00 -2.1277e+05 < 2.2e-16 ***
year_built1840                    2.8993e+04  1.3322e+00  2.1764e+04 < 2.2e-16 ***
 [ reached getOption("max.print") -- omitted 155 rows ]
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

Residual standard error: 1.963e+05 on 349122 degrees of freedom
Multiple R-squared: 0.2733,	Adjusted R-squared: 0.2726
> #R-squared value is 25.63%
> 
> # Perform predictions on test set
> reg_pred1 <- predict(fit1, as.matrix(x_test))
> 
> # See the comparison
> results <- data.frame(cbind(y_test, reg_pred1))
> results$diff <- results$reg_pred1 - results$y_test
> results$diff_squared <- '^' (results$diff, 2)
> # RMS = 
> rms = sqrt(mean(results$diff_squared))
> rms
[1] 198239.1
> 
> # RMS Error = 
> rms_error = rms/mean(results$y_test)
> rms_error
[1] 0.4358773
```
</details>
