---
title: "NYC Crime and Real Estate Data Project"
subtitle: "Final Tutorial - ENGM 182"
author: Vikhyat Khare, Omkar Kshirsagar, Carter Noordsij, John Sullivan
date: June 9, 2020
output:
  html_document:
    number_sections: true
---

# Server

Server is the part of shiny application where processing and updating of data, map, predicted values happen. We'll gradually build the code for Sever in steps. The inputIds in ui.R script will be linked in Sever script.


## Importing necessary libraries

```{r eval = FALSE}

library(shiny)
library(leaflet)
library(sf)
library(RColorBrewer)
library(zoo)
library(scales)
library(ggplot2)
library(caret)
library(randomForest)

```
## Loading the prediction model files

We'll be using Neural net and Random forest models to compute prices of real estate based on 4 inputs from user.

```{r eval = FALSE}

load("neural_net.rda")
load("random_forest.rda")

```
## Creating reactive merged dataframe and mapping
We start coding inside shinyServer function. In next steps, we build chunks of code that will go inside shinyServer function. They execute different processes in the app. Final code of server script is given in the end.

In this first step, we create a reactive obeject that changes the merged dateframe as per the user inputs: time selected and property type selected. The merged dataframe has spatial data. renderLeaflet function creates the map and shows NYC as we have put the coordinates.

```{r eval = FALSE}

shinyServer(function(input, output) {

    df <- reactive({
        if (housing_data_select()) {
            data = grouped_housing %>% 
                filter(category == input$prop_category, month_char == input$date_select) %>%
                select(zip_code, disp_data = !!input$data_select)
        } else {
            data = crime_scores %>%
                filter(month_char == input$date_select) %>%
                select(zip_code, disp_data = !!input$data_select)
        }
        tmp = merge(zip_sf, data, by.x="postalcode", by.y="zip_code", all.x=TRUE, no.dups=FALSE)
        return(tmp)
    })
    
    output$map <- renderLeaflet({
        leaflet() %>%
            addTiles() %>%
            setView(-74.0060,40.7128, zoom=11)
    })
    
})

```

## Mapping the reactive data and adding legend

Now, we add a color palette. Using reverse geocoding, we map the reactive data based on the ZIP code as per the attribute chosen by the user with radioButton and time chosen on sliderInput. We also highlight the ZIP code areas for mouse hover.

```{r eval = FALSE}
    
    observe({
        
        
        tmp = df()
        legend_labels = NULL
        
        pal = colorNumeric(
            palette = rev(brewer.pal(n=9, name = "RdYlGn")),
            domain = NULL
        )
        
        if (all(is.na(tmp$disp_data))) {
            func = labelFormat()
            tmp$disp_data = as.factor(rep("NA", nrow(tmp)))
            pal = colorFactor(palette = "#808080", domain = NULL)
        } else if (input$data_select == "avg_price_per_sqft") {
            func = labelFormat(prefix = " $")
        } else if (input$data_select == "total_proceeds") {
            func = labelFormat(prefix = " $", suffix = "M", transform=function(x) x/1E6)
        } else {
            func = labelFormat(prefix = " ")
        }
        
        leafletProxy("map", data = tmp) %>%
            clearControls() %>%
            clearShapes() %>%
            clearPopups() %>%
            addPolygons(
                fillColor = ~pal(disp_data),
                color = "#b2aeae", # you need to use hex colors
                fillOpacity = 0.7, 
                weight = 1, 
                smoothFactor = 0.2,
                # Highlight neighbourhoods upon mouseover
                highlight = highlightOptions(
                    weight = 3,
                    color = "black",
                    opacity = 1.0
                ),
                layerId = ~shape_id
            ) %>%
            addLegend(
                title=names(which(radioButtonOptions == input$data_select)),
                pal = pal, 
                values = ~disp_data, 
                position = "bottomright", 
                labFormat = func,
                labels = legend_labels
            )
    })
    

```
## Clicking on ZIP code areas

We create a function observeEvent that will check the click on a ZIP code area. It will show a pop up inside which we can see a graph of the chosen attribute over time. Code for the pop up in shown in the next step.

```{r eval = FALSE}
    # Observe shape clicks to display popup
    observeEvent(input$map_shape_click,{
        event = input$map_shape_click
        if (is.null(event))
            return()
        else {
            zip_clicked = zip_sf$postalcode[zip_sf$shape_id == event$id]
            curr = FALSE
            if (housing_data_select()) {
                data = grouped_housing %>%
                    filter(category == input$prop_category, zip_code == zip_clicked) %>%
                    select(month_char, !!input$data_select)
                title_suffix_str = paste("-", input$prop_category)
                if (input$prop_category == "All") {
                    title_suffix_str = paste(title_suffix_str, "Categories")
                }
                if (input$data_select %in% c("avg_price_per_sqft", "total_proceeds")) {
                    curr = TRUE
                }
            } else {
                data = crime_scores %>%
                    filter(zip_code == zip_clicked) %>%
                    select(month_char, weight_normalized)
                title_suffix_str = "over Time"
            }
            showPopup(data, zip_clicked, event$id, event$lat, event$lng, title_suffix_str, curr)
        }
    })
```

## Adding pop up

In this step, we add pop up on each ZIP code area showing the graph of the attribute chosen over time. Pop us also shows demographic information from the 2015 census.

```{r eval = FALSE}
showPopup <- function(df, zip, id, lat, lng, suff_str, curr = FALSE) {
        
        # Get plot object and other popup data for the chosen zip code
        zip_data = zip_sf[which(zip_sf$postalcode == zip),]
        if (nrow(df) > 0) {
            y = pull(df, 2)
            x = as.yearmon(pull(df, 1))
            title = paste(names(which(radioButtonOptions == colnames(df)[2])), suff_str)
            ylabel = names(which(radioButtonOptions == colnames(df)[2]))
            
            if (curr == TRUE) {
                if (max(y, na.rm=TRUE) > 1e6) {
                    label = label_number(prefix = "$", suffix = "M", scale = 1e-6)
                } else if (max(y, na.rm = TRUE) > 1e4) {
                    label =  label_number(prefix = "$", suffix = "K", scale = 1e-3)
                } else {
                    label = label_number(prefix = "$")
                }
            } else {
                label = label_number()
            }
            plot = ggplot(data=NULL, aes(x, y)) +
                geom_line() +
                geom_smooth() +
                labs(title = title, y = ylabel, x = "Month")  +
                scale_y_continuous(label = label, expand = c(0,0), limits = c(0, ifelse(colnames(df)[2] == "weight_normalized", 1, max(y, na.rm=TRUE)))) +
                theme_classic() +
                theme(plot.title = element_text(size = 11, hjust = 0.5), axis.title = element_text(size = 10))
    
            # Write svg file to temporary folder
            svg(filename= paste(folder,"plot.svg", sep = "/"),
                width = 500 * 0.01, height = 300 * 0.01)
            print(plot)
            dev.off()
    
            content <- paste(readLines(paste(folder,"plot.svg",sep="/")), collapse = "")
        } else {
            content = ""
        }
        
        if (is.na(zip_data$neighborhood)) {
            neighborhood_str = ""
        } else {
            neighborhood_str = paste(zip_data$neighborhood, " - ")
        }
        # Create popup
        leafletProxy("map") %>%
            addPopups(
                lng,
                lat,
                popup = paste0(
                    "<h4>", neighborhood_str, zip, "</h4><br/>",
                    content, "<br/>",
                    "<b>2015 Census Data</b><br/>",
                    "Per capita income:    $", round(zip_data$PerCapitaIncome),"</b><br/>",
                    "Total Population: ", zip_data$TotalPop, "</b><br/>",
                    "Unemployment rate: ", round(zip_data$Unemployed), "%"
                ),
                layerId = id,
                options = popupOptions(maxWidth = 500)
            )
    }
```
## Price predictor

In this step, we integrate the predictive models in the shiny app. It takes 4 inputs from the user: Age of the builing, area, sale year and ZIP code. ZIP code pulls demographic data from the 2015 cenus that's in the dataframe. It feeds these inputs in the model chosen by the user and gives a predicted price.

```{r eval = FALSE}
    
    price_text = eventReactive(input$predict_price, {
        if (input$zip != "")
            zip = unique_zips[which(unique_zips$select_names==input$zip), 1]
        else {
            zip = NA
        }
        input_data = data.frame(gross_square_feet = input$sqft, age = input$age, sale_year = input$sale_year, zip_code = zip, stringsAsFactors = FALSE)
        bad_cols = colnames(input_data)[colSums(is.na(input_data)) > 0]
        if (length(bad_cols) == 0){
            zip_data = zip_sf %>%
                filter(postalcode == zip) %>%
                slice(1) %>%
                data.frame() %>%
                select(postalcode, PerCapitaIncome, White, Black, Asian, Hispanic, Native, TotalPop.x = TotalPop, Unemployed, weight) %>%
                inner_join(input_data, by = c("postalcode" = "zip_code")) %>%
                select(-postalcode)
            if (is.na(zip_data)) {return(paste0("<font color=\"#eb4034\">Error: No valid census data for this zip code!</font>"))}
            if (input$model_select == "fit.rf") {
                predicted_val = predict(fit.rf, newdata=zip_data)
            } else if (input$model_select == "fit.nnet") {
                predicted_val = predict(fit.nnet, newdata=zip_data)
            } else if (input$model_select == "fit.bgcrt") {
                predicted_val = predict(fit.bgcrt, newdata=zip_data)
            }
            return(paste0("<br/>", dollar(predicted_val)))
        } else {
            msg = ""
            if ("gross_square_feet" %in% bad_cols) {
                msg = paste0(msg, "<br/>Error: Please enter a square footage!")
            }
            if ("age" %in% bad_cols) {
                msg = paste0(msg, "<br/>Error: Please enter building age!")
            }
            if ("sale_year" %in% bad_cols) {
                msg = paste0(msg, "<br/>Error: Please enter a sale year!")
            }
            if ("zip_code" %in% bad_cols) {
                msg = paste0(msg, "<br/>Error: Please enter a location!")
            }
            return(paste0("<font color=\"#eb4034\">", msg, "</font>"))
        }
    })

output$price = renderText({
        price_text()
    })

```
## Choice of attribute

housing_data_select takes a boolean value of either TRUE or FALSE depending on whether the attribute chosen by the user is within the choices provided.

```{r eval = FALSE}

outputOptions(output, "housing_data_select", suspendWhenHidden = FALSE)
    
    output$housing_data_select = reactive({
        return(housing_data_select())
    })
    
    housing_data_select = reactive({
        if (input$data_select %in% colnames(grouped_housing)[-c(1:3)]) {
            return(TRUE)
        } else {
            return(FALSE)
        }
    })
   
    
```
## Final code of server.R

Now we combine all chunks of code that we saw in steps so far inside the shinyServer function. This is the final script for server.R. This script along with ui.R and Global.R runs the shiny app.

``` {r eval = FALSE}

library(shiny)
library(leaflet)
library(sf)
library(RColorBrewer)
library(zoo)
library(scales)
library(ggplot2)
library(caret)
library(randomForest)

load("neural_net.rda")
# load(paste0(rda_loc, "bagged_cart.rda"))
load("random_forest.rda")

shinyServer(function(input, output) {
    
    output$housing_data_select = reactive({
        return(housing_data_select())
    })
    
    housing_data_select = reactive({
        if (input$data_select %in% colnames(grouped_housing)[-c(1:3)]) {
            return(TRUE)
        } else {
            return(FALSE)
        }
    })
   
    
    output$price = renderText({
        price_text()
    })
    
    price_text = eventReactive(input$predict_price, {
        if (input$zip != "")
            zip = unique_zips[which(unique_zips$select_names==input$zip), 1]
        else {
            zip = NA
        }
        input_data = data.frame(gross_square_feet = input$sqft, age = input$age, sale_year = input$sale_year, zip_code = zip, stringsAsFactors = FALSE)
        bad_cols = colnames(input_data)[colSums(is.na(input_data)) > 0]
        if (length(bad_cols) == 0){
            zip_data = zip_sf %>%
                filter(postalcode == zip) %>%
                slice(1) %>%
                data.frame() %>%
                select(postalcode, PerCapitaIncome, White, Black, Asian, Hispanic, Native, TotalPop.x = TotalPop, Unemployed, weight) %>%
                inner_join(input_data, by = c("postalcode" = "zip_code")) %>%
                select(-postalcode)
            if (is.na(zip_data)) {return(paste0("<font color=\"#eb4034\">Error: No valid census data for this zip code!</font>"))}
            if (input$model_select == "fit.rf") {
                predicted_val = predict(fit.rf, newdata=zip_data)
            } else if (input$model_select == "fit.nnet") {
                predicted_val = predict(fit.nnet, newdata=zip_data)
            } else if (input$model_select == "fit.bgcrt") {
                predicted_val = predict(fit.bgcrt, newdata=zip_data)
            }
            return(paste0("<br/>", dollar(predicted_val)))
        } else {
            msg = ""
            if ("gross_square_feet" %in% bad_cols) {
                msg = paste0(msg, "<br/>Error: Please enter a square footage!")
            }
            if ("age" %in% bad_cols) {
                msg = paste0(msg, "<br/>Error: Please enter building age!")
            }
            if ("sale_year" %in% bad_cols) {
                msg = paste0(msg, "<br/>Error: Please enter a sale year!")
            }
            if ("zip_code" %in% bad_cols) {
                msg = paste0(msg, "<br/>Error: Please enter a location!")
            }
            return(paste0("<font color=\"#eb4034\">", msg, "</font>"))
        }
    })

    df <- reactive({
        if (housing_data_select()) {
            data = grouped_housing %>% 
                filter(category == input$prop_category, month_char == input$date_select) %>%
                select(zip_code, disp_data = !!input$data_select)
        } else {
            data = crime_scores %>%
                filter(month_char == input$date_select) %>%
                select(zip_code, disp_data = !!input$data_select)
        }
        tmp = merge(zip_sf, data, by.x="postalcode", by.y="zip_code", all.x=TRUE, no.dups=FALSE)
        return(tmp)
    })
    
    output$map <- renderLeaflet({
        leaflet() %>%
            addTiles() %>%
            setView(-74.0060,40.7128, zoom=11)
    })
    
    observe({
        
        
        tmp = df()
        legend_labels = NULL
        
        pal = colorNumeric(
            palette = rev(brewer.pal(n=9, name = "RdYlGn")),
            domain = NULL
        )
        
        if (all(is.na(tmp$disp_data))) {
            func = labelFormat()
            tmp$disp_data = as.factor(rep("NA", nrow(tmp)))
            pal = colorFactor(palette = "#808080", domain = NULL)
        } else if (input$data_select == "avg_price_per_sqft") {
            func = labelFormat(prefix = " $")
        } else if (input$data_select == "total_proceeds") {
            func = labelFormat(prefix = " $", suffix = "M", transform=function(x) x/1E6)
        } else {
            func = labelFormat(prefix = " ")
        }
        
        leafletProxy("map", data = tmp) %>%
            clearControls() %>%
            clearShapes() %>%
            clearPopups() %>%
            addPolygons(
                fillColor = ~pal(disp_data),
                color = "#b2aeae", # you need to use hex colors
                fillOpacity = 0.7, 
                weight = 1, 
                smoothFactor = 0.2,
                # Highlight neighbourhoods upon mouseover
                highlight = highlightOptions(
                    weight = 3,
                    color = "black",
                    opacity = 1.0
                ),
                layerId = ~shape_id
            ) %>%
            addLegend(
                title=names(which(radioButtonOptions == input$data_select)),
                pal = pal, 
                values = ~disp_data, 
                position = "bottomright", 
                labFormat = func,
                labels = legend_labels
            )
    })
    
    
    # Observe shape clicks to display popup
    observeEvent(input$map_shape_click,{
        event = input$map_shape_click
        if (is.null(event))
            return()
        else {
            zip_clicked = zip_sf$postalcode[zip_sf$shape_id == event$id]
            curr = FALSE
            if (housing_data_select()) {
                data = grouped_housing %>%
                    filter(category == input$prop_category, zip_code == zip_clicked) %>%
                    select(month_char, !!input$data_select)
                title_suffix_str = paste("-", input$prop_category)
                if (input$prop_category == "All") {
                    title_suffix_str = paste(title_suffix_str, "Categories")
                }
                if (input$data_select %in% c("avg_price_per_sqft", "total_proceeds")) {
                    curr = TRUE
                }
            } else {
                data = crime_scores %>%
                    filter(zip_code == zip_clicked) %>%
                    select(month_char, weight_normalized)
                title_suffix_str = "over Time"
            }
            showPopup(data, zip_clicked, event$id, event$lat, event$lng, title_suffix_str, curr)
        }
    })
    
    showPopup <- function(df, zip, id, lat, lng, suff_str, curr = FALSE) {
        
        # Get plot object and other popup data for the chosen zip code
        zip_data = zip_sf[which(zip_sf$postalcode == zip),]
        if (nrow(df) > 0) {
            y = pull(df, 2)
            x = as.yearmon(pull(df, 1))
            title = paste(names(which(radioButtonOptions == colnames(df)[2])), suff_str)
            ylabel = names(which(radioButtonOptions == colnames(df)[2]))
            
            if (curr == TRUE) {
                if (max(y, na.rm=TRUE) > 1e6) {
                    label = label_number(prefix = "$", suffix = "M", scale = 1e-6)
                } else if (max(y, na.rm = TRUE) > 1e4) {
                    label =  label_number(prefix = "$", suffix = "K", scale = 1e-3)
                } else {
                    label = label_number(prefix = "$")
                }
            } else {
                label = label_number()
            }
            plot = ggplot(data=NULL, aes(x, y)) +
                geom_line() +
                geom_smooth() +
                labs(title = title, y = ylabel, x = "Month")  +
                scale_y_continuous(label = label, expand = c(0,0), limits = c(0, ifelse(colnames(df)[2] == "weight_normalized", 1, max(y, na.rm=TRUE)))) +
                theme_classic() +
                theme(plot.title = element_text(size = 11, hjust = 0.5), axis.title = element_text(size = 10))
    
            # Write svg file to temporary folder
            svg(filename= paste(folder,"plot.svg", sep = "/"),
                width = 500 * 0.01, height = 300 * 0.01)
            print(plot)
            dev.off()
    
            content <- paste(readLines(paste(folder,"plot.svg",sep="/")), collapse = "")
        } else {
            content = ""
        }
        
        if (is.na(zip_data$neighborhood)) {
            neighborhood_str = ""
        } else {
            neighborhood_str = paste(zip_data$neighborhood, " - ")
        }
        # Create popup
        leafletProxy("map") %>%
            addPopups(
                lng,
                lat,
                popup = paste0(
                    "<h4>", neighborhood_str, zip, "</h4><br/>",
                    content, "<br/>",
                    "<b>2015 Census Data</b><br/>",
                    "Per capita income:    $", round(zip_data$PerCapitaIncome),"</b><br/>",
                    "Total Population: ", zip_data$TotalPop, "</b><br/>",
                    "Unemployment rate: ", round(zip_data$Unemployed), "%"
                ),
                layerId = id,
                options = popupOptions(maxWidth = 500)
            )
    }
    
    outputOptions(output, "housing_data_select", suspendWhenHidden = FALSE)
})


```

